---
title: "R for data science tutorial"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    self_contained: true
    downcute_theme: "chaos"
---
```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(cache = TRUE)
```

# 3. Data visualization
## 3.1 Prerequisites

```{r}
library(tidyverse)
```

## 3.2
### 3.2.1 The mpg data frame
```{r}
mpg
```

### 3.2.2 Creating a ggplot
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```
### 3.2.4 Exercises
```{r}
# How many rows are in mpg? How many columns
dim(mpg)
```
```{r eval = FALSE}
# What does the drv variable describe?
?mpg
drv <- "the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd"
```
```{r}
# Make a scatterplot of hwy vs cyl
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))

# What happens if you make a scatterplot of class vs drv? Why is the plot not useful?
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))

```

## 3.3 Aesthetic mappings
> "The greatest value of a picture is when it forces us to notice what we never expected to see." - John Tukey

Map the colors of the points to the class variable to reveal the class of each car
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```
Mapping to the shapes of the points instead
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
# Additional groups beyond 6 will go unplotted when using the shape aesthetic
```
### 3.3.1 Exercises
```{r}
# Why are the points not blue?
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
## fix by bringing the color customization outside of the aes argument
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")

# Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs continuous variables?
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = fl, color = hwy))

# What happens if you map the same variable to multiple aesthetics?
ggplot(data = mpg) +
  geom_point(mapping = aes(x = cty, y = cty, color = cty))
```
```{r eval = FALSE}
# What does the stroke aesthetic do? What shapes does it work with?
?geom_point
```
``` {r}
## For shapes that have a border, the stroke aesthetic modifies the width of the border

# What happens if you map an aesthetic to something other than a variable name?
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = fl, color = displ < 5))
```
## 3.5 Facets
To facet your plot by a single variable, use facet_wrap()
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)
```
To facet your plot on the combination of two variables, add facet_grid()
to your plot call
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```
### 3.5.1 Exercises
```{r}
# What happens if you facet on a continuous variable?
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = class)) +
  facet_wrap(~ hwy)

# What plots does the following code make? What does . do?
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)

# What are the advantages/disadvantages to using faceting instead of the color aesthetic?
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```
## 3.6 Geometric objects
```{r}
# left
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))

# right
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

